Diversity and evolution of centromere repeats in the maize genome

### 2014-06-26 response.tex

\noindent Jeffrey Ross-Ibarra and Paul Bilinski

\section*{Editor:}

Personally I wonder whether genome size differences between maize landraces will correlate with differences in CentC content (p. 11), because it is known from old cytogenetic studies that they may vary in their number of non-centromeric heterochromatic knobs. \\

\res{We agree with the editor that the correlation between CentC abundance and knob number or genome size is of great interest. 
 Our current panel of maize lines available from the hapmap project is limited in its number of knobs, as many are inbred lines with relatively uniform knob abundance when compared to the variation seen in landraces, and therefore beleive it is not a suitable panel to investigate this question.
However, we are investigating the correlation in another project using a broad panel of approximately 300 maize landraces across an altitudinal gradient.  
Though preliminary, our as yet punpublished findings are that CentC exhibits an opposite trend to that observed in knobs; CentC increases when knob count and genome size decreases along an altitudinal gradient.  
We are currently in the process of further investigating this question, and hope to save these findings for another publication.}

@@

\section*{Reviewer 1:}

Of particular note is the finding that "virtually all the large arrays of CentC in the maize genome derive from one of the two ancestral genomes present in modern day Zea”.  This is an interesting observation.  However, there are several potential pitfalls to this analysis that have not been addressed.  First, the authors state that "because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length."  In fact, large tandem repeat clusters are much more likely to be misassembled than shorter counterparts.  \\

\res{We agree with the reviewer that large clusters are more likely to experience misassembly within the clusters.  
We had intended to refer to a lower likelihood of being in the wrong genomic location with regard to subgenome.  
Though large clusters are more likely to be incorrect in terms of copy number count or exact sequence, large clusters are more likely to be properly located on the maize iMap (now cited in the manuscript), a resource utilized in genome assembly that blended the physical, genetic, and optical maps of the maize genome.  
We have altered the text in the manuscript to properly orient our point.}

Also, tandem repeats are often broken up by retrotransposon insertions, so that large uninterrupted CentC runs could be interpreted quite differently as "immune to retrotransposon insertion" (although this is certainly not what I suggest is going on).

\res{The reviewer's point addressing novel insertions breaking up large clusters is very insightful.
To investigate this, we reanalyzed cluster size data allowing for variable distances between clusters and variable cluster size to determine membership.
The results are included in a new supplementary, and its consequences to further analyses such as subgenome counts are discussed later in our response.}

@@

This has led to a large number of incorrect "joins" involving repeats.  
For the same reason perfect tandem repeats would have been collapsed into a single sequence in the final maize genome assembly.  
Therefore, the large CentC repeats are almost certainly misassembled and, if correct, would represent not the largest arrays but rather the most degenerate arrays of CentC repeats.\\

\res{The reviewer makes an excellent point that the most degenerate arrays are also the most likely to not be collapsed into single CentC repeats during automated assembly.
Indeed, automated assembly may explain the apparent lack of identical tandem duplicates across the genome and, more importantly, directly impact our ability to determine copy number in extremely large clusters.  
However, we would expect that any bias in assembly would impact clusters from the different subgenomes similarly and therefore should not impact our overall findings regarding subgenome.  }

@@

\res{Furthermore, our ability to judge subgenome of origin is not likely to be impacted through assembly issues. In most cases, the origin of CentC clusters were determined with adjacent single copy markers specific to one of the subgenome as outlined in Schnable et al 2011.  The number of CentC's in the cluster would not change the origin call.}

\res{Regarding diversity of CentC copies from chromosomes with automated assembly, we also do not see striking differences.  
Centromeres 2 and 5 were carefully assembled BAC-by-BAC as part of the reference genome (see Wolfgruber et al. 2009).  CentC in these centromeres  show similar levels of diversity across CentC repeats compared to other chromosomes.  
We believe that the relatively uniform level of diversity observed across CentC is likely due to mutation over long periods of time.  
Though assembly errors likely impact our ability to detect many of the tandem duplication instances of CentC in the maize genome, we believe that identical tandem duplicates account for a relatively small quantity of overall CentC copies as evidenced by centromeres 2 and 5. }

@@

In the abstract, the authors are even bolder with the following statement: "Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that the vast majority of all CentC repeats derive from one of the parental genomes.”\\

\res{As we explain above, we believe our ability to determine subgenome of origin is not comprised through assembly errors, and our reanalysis considering clusters of varying size and distance supports our original idea.
The signal is likely driven by the strong bias of CentC from subgenome, showing that regardless of our definition of tandem clusters, CentC membership is biased to subgenome 1.
We have included a new supplementary table in the text to report these findings, and have altered the text to include the reanalysis.}

@@

Wild maize relatives (other subspecies) have the same five centromeres derived from each of the two maize ancestral sub genomes as the reference inbred genome, yet the authors themselves show by FISH that ALL centromeres of these wild species have similarly high amounts of CentC.\\

\res{We do acknowledge that FISH signal does indicate higher copy numbers of CentC in wild Zea across all chromosomes, but examination of the signals does not suggest uniformity across chromosomes.  
For example, in figure 4B, the fluorescent signal on parviglumis chromosome 2 is much stronger than chromosome 3.  
The centromeres on these chromosomes originate from different subgenomes.  
Furthermore, we believe a qualitative measure such as fluorescence is not a good measure for precise CentC content and are hesitant to try to compare CentC content quantitatively using FISH alone. }

@@

I do agree with the author's finding that proximal CentC copies are more similar than distal ones, and that "The tandem nature of CentC suggests it increases in copy number through local duplications that produce initially identical copies." Similar conclusions were found by Ma and Jackson (\url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC1361721/}).\\

\res{We thank the reviewer for pointing out this paper to us and have incorporated the citation.  We're glad that our finding in maize that uses a different method is consistent with standing hypotheses.}

@@

However, the authors observe that pairs of  CentC on different chromosomes share very high sequence similarity.  This again is a very interesting observation if one could be sure that the physical map is correct, but at least the early versions of the B73 map were fraught with major errors (entire BAC contigs and their associated centromeres being anchored to the wrong chromosome - see Schnable genome paper).  How sure are the authors that the specific repeats they find on multiple chromosomes are truly present, as opposed to having been misassigned to the wrong chromosome? This point should probably be addressed in the manuscript.\\

\res{We acknowledge that large assembly errors in regards to the exact location on a chromosome likely exist in the maize reference genome (for example a cluster of CentC on the tip of the short arm of chromosome10 that belongs on chromosome 7), and have added text to this effect in the discussion.  
Nonetheless, we suspect our result is robust to these errors.
Our main finding is that CentCs can be most closely related to CentCs from different clusters.
Because most clusters are on different BACs in the genome assembly, our result would still provide evidence of homoplasy even if BACs were improperly assigned to a chromosome. 
While some of the highly-related pairs that we observe on different chromosomes may be due to assembly error (i.e. the BAC is placed on the wrong chromosome) we suspect it is unlikely that all such pairs can be explained by miss-assembly.  
Furthermore, looking only at the fully sequenced centromeres 2 and 5, we find several pairs of CentC on different chromosomes that are reciprocally closest genetically compared to all other CentC in the genome.  
We have included text in the manuscript to reflect this more robust finding. 
It is also worth noting that centromeric BAC assemblies have improved since v1 of the reference genome, and LD-mapping with GBS data (unpublished) shows that while many BACs are in the wrong order, few of the BACs physically placed in a centromeres are on the wrong chromosome. }

@@

\section*{Reviewer 2:}

1) Authors thought tandem duplication is the most important mechanism for the evolution of the repeat. However, as authors pointed out, tandem duplication generates similar sequence in near-by copies. I think gene conversion and unequal crossing over should play important roles for repeat evolution (in some part). For example, in Arabidopsis species (eg, Kawabe and Charlesworth 2007 JME), chimeric or mosaic structure repeats were found which could be originated by gene conversion or unequal crossing over. I think it is necessary to discuss about gene conversion and unequal crossing over in this manuscript.\\

\res{We consider unequal crossing over as a mechanism of tandem duplication; this is now clarified in the text.
Gene conversion likely serves to homogenize repeats across clusters and chromosomes, and may help explain why related sequences are not always physically close.
Nonetheless, most of our clusters contain many unique CentC, suggesting either that gene conversion is infreqeunt or point mutations are common. 
We have added text to the discussion mentioning gene conversion.}

@@

2) Although it is not the main part in this manuscript, I wonder which arrays of CentC had decreased (or disappeared) from domesticated maize genome. Because not all CentC arrays form kinetochores, there is different importance of arrays. If it is possible to distinguish kinetochore forming and other arrays, please show level of decrements between them.\\

\res{We are only able to analyze genetic diversity of CentC copies within active/inactive kinetochores with confidence in the fully sequenced centromeres 2 and 5.  
On these chromosomes, we find no difference in genetic diversity of arrays inside and outside the active kinetochores.  
The largest array on chromosome 2 is found in the kinetochore, but it does not show higher uniformity when compared to the arrays on chromosome 5, none of which are found in the kinetochore.  
We have added text in the manuscript to emphasize these findings.
Unfortunately, it is not possible to determine whether active or inactive arrays are responsible for the change in abundance observed during domestication.}

3) Authors conclude importance of tandem duplication because of the close location of highest similarity pairs and homogenization within arrays. However authors pointed out some distances between closest repeat and suggest that tandem duplication could be RELATIVELY old event. It is possible to estimate divergence between such pairs to show rough estimate of divergence times.\\

\res{The short nature of individual repeats provides for a small mutational target, causing standard estimates of age to have high variance.  
Standard estimation of age also assumes fixation, while many CentC variants likely exist as copy number variants in the population. 
More sophisticated analysis (e.g. Thornton 2007 Genetics) require knowledge of the frequency of the copy number variant, which we do not have.
Furthermore, we show in the text that homoplasious mutation is a problem, making it difficult to discern ancestry of CentC copies without error.  
As such, we felt that age estimation would be complicated and unlikely to be very accurate.}

### 2014-06-26 centc.tex
edits done, needs a final read through

We found a total of 12,162 CentC copies in the maize reference genome and unassembled BACs.  
Of these, 8,259 were unique over their full length.
While centromeres 2 and 5 are the only chromosomes with high confidence sequencing of CentC copies, the levels of diversity observed on these two chromosomes is comparable to the rest of the genome, suggesting that current assemblies of CentC sequence may estimate sequence variation with some accuracy.
No CentC sequence occurred more than 10 times in the genome, and the vast majority ($>$75\%, Table \ref{supp.occurence}) of non-unique CentC variants occurred only twice.  
Of the 2,266 non-unique CentC sequences, only 3 were tandem, identical duplicates.  
%Excluding the 1,523 copies that occur on the unassembled BACs, 
Genome-wide CentC locations also show that nearly all of the 10,639 CentC copies on chromosomes 1-10 are found in clusters; only 14 occurred as solo copies.  
Clusters varied in width from single CentC copies to 84 Kb with a mean of $\sim$7 Kb ($\sim$45 CentC copies). 
Chromosomes varied greatly in CentC copy number, though we know that centromere assemblies for all of the chromosomes are not complete.   
For example, CENH3-ChIP sequence from an oat-maize addition line with one maize chromosome \citep{kynast2001} has many reads that map to the unassembled BACs (Table \ref{supp.gaby}).  
In particular, chromosome 6 has many more reads aligning to the unassembled BACs than it did to its own centromere repeats, suggesting a particularly incomplete assembly.  
Examining total repeat number, chromosome 7 had the most CentC, with 3,200 copies, while chromosome 6 had the fewest with 32 copies.

@@

We then identified how many large clusters of CentC were retained from each of the two parental genomes that comprise the extant maize genome, referred to here as subgenome 1 and subgenome 2 (Figure \ref{circos}).  
Previous work identified the parental genome for individual chromosomal segments \citep{Schnable2011} and centromeres \citep{Wang2012}.  
Due to the maize iMap resource \citep{zhou2009single}, large clusters are less likely to be misplaced within the genome.
Therefore, we focused our analyses on the 52 clusters $>$10 Kb in length (Supplementary Fig \ref{Supp_clusterhist}).
We assign clusters to a subgenome if they are flanked by two regions identified as originating from the same  subgenome.  
Thirty-eight of these clusters could be assigned to subgenome 1 (out of 43 assignable).
If we restrict the analysis to clusters  $>$20 Kb with clear assignment to one subgenome, all 16 clusters were found in subgenome 1 regions.  
Even correcting for the genome-wide overrepresentation of subgenome 1 (64.7\% of assigned base pairs), these results suggest a strong inequality in the origin of large CentC clusters (Fisher's exact test, $p<0.005$ for both 10 Kb and 20 Kb clusters).
One cluster $>$20 Kb falls within an unassigned region on chromosome 3.
Our results are robust to varying definitions of clusters, as the bias is maintained when allowing for varying distances to define clusters \ref{supp.vartandem}.

@@

The tandem nature of CentC suggests it increases in copy number through local duplications that produce initially identical copies.  
Tandem duplications can occur through a variety of means, including slippage of the DNA polymerase \citep{jeeeeeeff.} or recombination that could lead to unequal crossing over or gene conversion \citep{edlund1981recombination}.
Tandem duplication predicts that  clusters of CentC should be more closely related than CentC from different clusters.  
Comparisons of genetic and physical distance among CentC repeats on chromosomes 2 and 5 shows that average genetic similarity is highest within clusters (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10-50 Kb (Figure \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
Interestingly, we did not observe differences in genetic diversity when considering whether an array was found within or outside of the active kinetochore.

@@

The decreased genetic distance among CentCs in local clusters on chromosome 2 and 5 suggest that many of the genetic groupings discovered in our genome-wide analysis should correspond to local clusters of repeats. 
However, repeats within individual clusters are frequently found in different genetic groups as defined by principle coordinate analysis (Figure \ref{pcoa}).
Several of these best pairs on different chromosomes are found between chromosomes 2 and 5, suggesting the phenomenon is not an artifact of errors in assembly.
A comparison of shared mutations across all pairs of CentC sequences reveals a potential explanation.  
Of the $\sim 74$ million possible pairs,  approximately 6 million share $\geq 2$ mutations different from the genome-wide consensus, causing CentC copies to group with sequences that share mutations irrespective of their physical distance.  
Comparing several triplets at random from our alignment confirms that two sequences in one PCoA assignment share greater pairwise identity than two sequences adjacent to one another in different PCoA groups.  
A simple forward simulation (see Methods) suggests this pattern could be due entirely to homoplasy rather than long-distance movement of CentC repeats.  
By stochastically applying mutations to an initially homogeneous group of repeat sequences, we find that plausible parameter values produce $\sim 10$ million pairs of repeats sharing $\geq 2$ mutations.

@@

Shotgun sequence data from the maize HapMap v2 \citep{Chia2012}, reveals a significantly greater abundance of CentC in teosinte than in domesticated maize (p$<0.01$; Figure \ref{abundance}).   
Further support for differences between maize and its wild relatives comes from shotgun sequence data from \emph{Z. luxurians} \citep{Tenaillon2011}.  
Analysis of these data find nearly twice as much CentC in \emph{Z. luxurians} as the maize inbred B73.  
To corroborate these results, we performed fluorescent in-situ hybridization of F1 crosses between inbred maize and teosinte to determine if cytological observations agreed with our sequencing findings.   
FISH data supports our observation that the teosintes \emph{parviglumis} and \emph{Z. luxurians} have more CentC than inbred maize (Figure \ref{abundance}).
Using whole genome shotgun PacBio long reads, we further investigated the overall structure of repeats across the different \emph{Zea} species.
Percentages of the libraries showing tandem repeats did not differ greatly across the taxa (Figure \ref{supp.pacbio}).

@@

Our analysis of centromere repeat diversity across the maize genome identifies thousands of copies exhibiting tremendous diversity. 
But while we can cluster the repeats into groups of related sequences, these groups have little relation to current or ancient maize chromosomes (Figure \ref{pcoa}).
We find no evidence of chromosome specific repeats as observed in \emph{Arabidopsis} species \citep{Kawabe2005, Pontes2004}, suggesting the presence of a mechanism that homogenizes repeats across centromeres on different chromosomes.
%jeff, this needs to go here, but damnit its terrible.  i can't argue this well since i feel like its such a dumb point.
Assembly errors, such as misplaced BACs known to occur in the maize genome or problems with assembling capturing biologically real CentC diversity, may also influence our ability to discover chromosomal patterns, and will only be improved over time with better sequencing technologies.
We further verify that Cent4, once thought to be a chromosome-specific centromere repeat \citep{Page2001}, appears to be a poorly characterized tandem repeat or nonautonomous retroelement, but is not associated with the centromere, though it is present in much of modern maize and absent in the teosintes \citep{albert2010}.  
Inter-chromosomal transposition was postulated by \citet{Shi2010} as an explanation for marker genotypes showing evidence of recombination, and such transposition could also explain the similarity of repeats across all chromosomes.

Additionally, we find that virtually all the large arrays of CentC in the maize  genome derive from one of the two ancestral genomes present in modern day \emph{Zea} (Figure \ref{circos}).  
One possible explanation for the biased inheritance pattern is that centromere inactivation mirrors the differential expression seen between the subgenomes.  
Centromere inactivation could then lead to differential loss of CentC clusters, such that large clusters remain only in the active centromeres.
%POTENTIAL INSERT: We do not observe differences in genetic distances between CentC clusters on chromosomes 2 and 5 that are inside or outside the active kinetochore, which may suggest that purges occur on a large scale versus sustained mutation over an extended time.
Large CentC clusters are biased in favor of subgenome 1 \ref{supp.vartandem}, the subgenome known to have lower gene loss and higher average expression than maize subgenome 2 \citep{Schnable2011}. 
Consistent with a correlation between expression and CentC loss in subgenome 2, \citet{ReinhartBartel2002} found evidence of regulatory microRNA’s corresponding to centromeric repeats.
An alternative explanation is that the subgenome 1 parent had more or larger CentC arrays, leading to an initial unequal representation in the ancient polyploid ancestor of maize.

@@

One unusual result from our sequence comparison was the finding that pairs of CentC on different chromosomes share very high sequence similarity.  
Our simulations suggest that, under realistic assumptions about mutation rate and divergence time, such a pattern is possible due to homoplasious mutation alone.  
Roughly 80\% of the CentC repeats have their closest genetic relative on the same chromosome, as expected under a model of tandem duplication, but only 14\% of closest genetic pairs are found within 10 Kb of each other.
Though assembly errors may explain a portion of these relationships, we find several closest distances pairs between chromosomes 2 and 5, suggesting our observation is biologically real.
We speculate that the vast majority of  CentC’s in the genome are thus a result of relatively old tandem duplications, and that sufficient time has occurred since duplication for rearrangements and mutations to break up patterns of identical tandem repeats.

@@

\clearpage
\begin{table}[h!]
\label{supp.vartandem} 
\caption{Number of clusters in each subgenome under varying criteria for distances apart.  Minimum cluster size is 10,000 bp, though trends remain the same under different mininum cluster sizes.}\label{supp.adjustclust} 
% For LaTeX tables use
\begin{tabular}{ccccccc}
\hline\noalign{\smallskip}
Distance ( bp) &  Unassigned  bp &  Clusters Unassigned &  Subgenome1  bp &  Clusters Subgenome1 & Subgenome2  bp &  Clusters Subgenome2 \\\noalign{\smallskip}\hline\noalign{\smallskip}				
1000 & 96504 & 6 & 829419 & 39 & 92126 & 7\\
2000 & 109482 & 7 & 925028 & 43 & 115538 & 7\\
3000 & 128492 & 8 & 969892 & 44 & 117094 & 7\\
4000 & 128492 & 8 & 994104 & 41 & 124020 & 7\\
5000 & 128492 & 8 & 1026845 & 40 & 136025 & 6\\
6000 & 134203 & 8 & 1054805 & 39 & 140679 & 6\\
7000 & 134203 & 8 & 1069516 & 38 & 140679 & 6\\
8000 & 134203 & 7 & 1139270 & 36 & 140679 & 4\\
9000 & 142786 & 7 & 1162918 & 37 & 147864 & 4\\
10000 & 142786 & 7 & 1179529 & 36 & 149125 & 4\\
11000 & 144185 & 7 & 1198451 & 36 & 149125 & 4\\
12000 & 144185 & 6 & 1203008 & 34 & 149125 & 4\\
13000 & 145273 & 6 & 1210187 & 32 & 149125 & 4\\
14000 & 145273 & 6 & 1210329 & 31 & 149125 & 4\\
15000 & 145273 & 6 & 1215760 & 30 & 149125 & 4\\
16000 & 145273 & 5 & 1224106 & 27 & 152385 & 4\\
17000 & 145273 & 5 & 1224106 & 25 & 152385 & 4\\
18000 & 146060 & 5 & 1224106 & 25 & 157135 & 4\\
19000 & 152703 & 5 & 1229229 & 25 & 157135 & 4\\
20000 & 163555 & 6 & 1234836 & 25 & 157135 & 3\\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

>>>>

We found a total of 12,162 CentC copies in the maize reference genome and unassembled BACs.  
Of these, 8,259 were unique over their full length.
While centromeres 2 and 5 are the only chromosomes with high confidence sequencing of CentC copies, the levels of diversity observed on these two chromosomes is comparable to the rest of the genome (data not shown), suggesting that current assemblies of CentC sequence may estimate sequence variation with some accuracy.
No CentC sequence occurred more than 10 times in the genome, and the vast majority ($>$75\%, Table \ref{supp.occurence}) of non-unique CentC variants occurred only twice.  
Of the 2,266 non-unique CentC sequences, only 3 were tandem, identical duplicates.  
%Excluding the 1,523 copies that occur on the unassembled BACs, 
Genome-wide CentC locations also show that nearly all of the 10,639 CentC copies on chromosomes 1-10 are found in clusters; only 14 occurred as solo copies.  
Clusters varied in width from single CentC copies to 84 Kb with a mean of $\sim$7 Kb ($\sim$45 CentC copies). 
Chromosomes varied greatly in CentC copy number, though we know that centromere assemblies for all of the chromosomes are not complete.   
For example, CENH3-ChIP sequence from an oat-maize addition line with one maize chromosome \citep{kynast2001} has many reads that map to the unassembled BACs (Table \ref{supp.gaby}).  
In particular, chromosome 6 has many more reads aligning to the unassembled BACs than it did to its own centromere repeats, suggesting a particularly incomplete assembly.  
Examining total repeat number, chromosome 7 had the most CentC, with 3,200 copies, while chromosome 6 had the fewest with 32 copies.

@@

We then identified how many large clusters of CentC were retained from each of the two parental genomes that comprise the extant maize genome, referred to here as subgenome 1 and subgenome 2 (Figure \ref{circos}).  
Previous work identified the parental genome for individual chromosomal segments \citep{Schnable2011} and centromeres \citep{Wang2012}.  
Due to the maize iMap resource \citep{zhou2009single}, large clusters are less likely to be misplaced within the genome.
Therefore, we focused our analyses on the 52 clusters $>$10 Kb in length (Supplementary Fig \ref{Supp_clusterhist}).
We assign clusters to a subgenome if they are flanked by two regions identified as originating from the same  subgenome.  
Thirty-eight of these clusters could be assigned to subgenome 1 (out of 43 assignable).
If we restrict the analysis to clusters  $>$20 Kb with clear assignment to one subgenome, all 16 clusters were found in subgenome 1 regions.  
Even correcting for the genome-wide overrepresentation of subgenome 1 (64.7\% of assigned base pairs), these results suggest a strong inequality in the origin of large CentC clusters (Fisher's exact test, $p<0.005$ for both 10 Kb and 20 Kb clusters).
One cluster $>$20 Kb falls within an unassigned region on chromosome 3.
This difference between the subgenomes is robust to different criteria of cluster size and distance (Table \ref{tab:suppvartandem}).

@@

The tandem nature of CentC suggests it increases in copy number through local duplications that produce initially identical copies.  
Tandem duplications can occur through a variety of means, including slippage of the DNA polymerase \citep{jeeeeeeff.} or recombination that could lead to unequal crossing over or gene conversion \citep{edlund1981recombination}.
Tandem duplication predicts that  clusters of CentC should be more closely related than CentC from different clusters.  
Comparisons of genetic and physical distance among CentC repeats on chromosomes 2 and 5 shows that average genetic similarity is highest within clusters (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10-50 Kb (Figure \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).

@@

The decreased genetic distance among CentCs in local clusters on chromosome 2 and 5 suggest that many of the genetic groupings discovered in our genome-wide analysis should correspond to local clusters of repeats. 
However, repeats within individual clusters are frequently found in different genetic groups as defined by principle coordinate analysis (Figure \ref{pcoa}).
We observed multiple pairs of CentC which occur in the same genetic cluster in spite of being on separated on the completely sequenced centromeres 2 and 5, suggesting that our result is not simply an artifact of errors in assembly.
A comparison of shared mutations across all pairs of CentC sequences reveals a potential explanation.  
Of the $\sim 74$ million possible pairs,  approximately 6 million share $\geq 2$ mutations different from the genome-wide consensus, causing CentC copies to group with sequences that share mutations irrespective of their physical distance.  
Comparing several triplets at random from our alignment confirms that two sequences in one PCoA assignment share greater pairwise identity than two sequences adjacent to one another in different PCoA groups.  
A simple forward simulation (see Methods) suggests this pattern could be due entirely to homoplasy rather than long-distance movement of CentC repeats.  
By stochastically applying mutations to an initially homogeneous group of repeat sequences, we find that plausible parameter values produce $\sim 10$ million pairs of repeats sharing $\geq 2$ mutations.

@@

Shotgun sequence data from the maize HapMap v2 \citep{Chia2012}, reveals a significantly greater abundance of CentC in teosinte than in domesticated maize (p$<0.01$; Figure \ref{abundance}).   
Further support for differences between maize and its wild relatives comes from additional  sequence  from \emph{Z. luxurians} \citep{Tenaillon2011}.  
Analysis of these data find nearly twice as much CentC in \emph{Z. luxurians} as the maize inbred B73.  
To corroborate these results, we performed fluorescent in-situ hybridization of F1 crosses between inbred maize and teosinte to determine if cytological observations agreed with our sequencing findings.   
FISH data supports our observation that the teosintes \emph{parviglumis} and \emph{Z. luxurians} have more CentC than inbred maize (Figure \ref{abundance}).
Using whole genome shotgun PacBio long reads, we further investigated the overall structure of repeats across the different \emph{Zea} species.
Percentages of the libraries showing tandem repeats did not differ greatly across the taxa (Figure \ref{supp.pacbio}).

@@

Our analysis of centromere repeat diversity across the maize genome identifies thousands of copies exhibiting tremendous diversity. 
But while we can cluster the repeats into groups of related sequences, these groups have little relation to current or ancient maize chromosomes (Figure \ref{pcoa}).
We find no evidence of chromosome specific repeats as observed in \emph{Arabidopsis} species \citep{Kawabe2005, Pontes2004}, suggesting the presence of a mechanism that homogenizes repeats across centromeres on different chromosomes.
Although we believe our results relatively robust to assembly errors, misplaced BACs and collapsed tandem copies almost certainly occur in the maize  reference genome and may influence our assessment of chromsome-wide patterns.
We further verify that Cent4, once thought to be a chromosome-specific centromere repeat \citep{Page2001}, appears to be a poorly characterized tandem repeat or nonautonomous retroelement, but is not associated with the centromere, though it is present in much of modern maize and mostly absent in the teosintes \citep{albert2010}.  
Inter-chromosomal transposition was postulated by \citet{Shi2010} as an explanation for marker genotypes showing evidence of recombination, and such transposition could also explain the similarity of repeats across all chromosomes.

Additionally, we find that virtually all the large arrays of CentC in the maize  genome derive from one of the two ancestral genomes present in modern day \emph{Zea} (Figure \ref{circos}).  
One possible explanation for the biased inheritance pattern is that centromere inactivation mirrors the differential expression seen between the subgenomes.  
Centromere inactivation could then lead to differential loss of CentC clusters, such that large clusters remain only in the active centromeres.
We do not observe differences in genetic distances between CentC clusters on chromosomes 2 and 5 that are inside or outside the active kinetochore, which may suggest that purges occur on a large scale versus sustained mutation over an extended time.
Large CentC clusters are biased in favor of subgenome 1 \ref{tab:suppvartandem}, the subgenome known to have lower gene loss and higher average expression than maize subgenome 2 \citep{Schnable2011}. 
Consistent with a correlation between expression and CentC loss in subgenome 2, \citet{ReinhartBartel2002} found evidence of regulatory microRNA’s corresponding to centromeric repeats.
An alternative explanation is that the subgenome 1 parent had more or larger CentC arrays, leading to an initial unequal representation in the ancient polyploid ancestor of maize.

@@

One unusual result from our sequence comparison was the finding that pairs of CentC on different chromosomes share very high sequence similarity.  
Our simulations suggest that, under realistic assumptions about mutation rate and divergence time, such a pattern is possible due to homoplasious mutation alone.  
Roughly 80\% of the CentC repeats have their closest genetic relative on the same chromosome, as expected under a model of tandem duplication, but only 14\% of closest genetic pairs are found within 10 Kb of each other.
Though assembly errors may explain a portion of these relationships, we find several closest genetic pairs separated in the fully sequenced centromeres of chromosomes 2 and 5, suggesting our observation is biologically real.
We speculate that the vast majority of  CentC’s in the genome are thus a result of relatively old tandem duplications, and that sufficient time has occurred since duplication for rearrangements and mutations to break up patterns of identical tandem repeats.

@@

\clearpage
\begin{table}[h!]
\caption{Number of clusters in each subgenome under varying criteria for distances apart.  Minimum cluster size is 10,000 bp, though trends remain the same under different mininum cluster sizes.}\label{supp.adjustclust} 
% For LaTeX tables use
\begin{tabular}{ccccccc}
\hline\noalign{\smallskip}
Distance ( bp) &  Unassigned  bp &  Clusters Unassigned &  Subgenome1  bp &  Clusters Subgenome1 & Subgenome2  bp &  Clusters Subgenome2 \\\noalign{\smallskip}\hline\noalign{\smallskip}				
1000 & 96504 & 6 & 829419 & 39 & 92126 & 7\\
2000 & 109482 & 7 & 925028 & 43 & 115538 & 7\\
3000 & 128492 & 8 & 969892 & 44 & 117094 & 7\\
4000 & 128492 & 8 & 994104 & 41 & 124020 & 7\\
5000 & 128492 & 8 & 1026845 & 40 & 136025 & 6\\
6000 & 134203 & 8 & 1054805 & 39 & 140679 & 6\\
7000 & 134203 & 8 & 1069516 & 38 & 140679 & 6\\
8000 & 134203 & 7 & 1139270 & 36 & 140679 & 4\\
9000 & 142786 & 7 & 1162918 & 37 & 147864 & 4\\
10000 & 142786 & 7 & 1179529 & 36 & 149125 & 4\\
11000 & 144185 & 7 & 1198451 & 36 & 149125 & 4\\
12000 & 144185 & 6 & 1203008 & 34 & 149125 & 4\\
13000 & 145273 & 6 & 1210187 & 32 & 149125 & 4\\
14000 & 145273 & 6 & 1210329 & 31 & 149125 & 4\\
15000 & 145273 & 6 & 1215760 & 30 & 149125 & 4\\
16000 & 145273 & 5 & 1224106 & 27 & 152385 & 4\\
17000 & 145273 & 5 & 1224106 & 25 & 152385 & 4\\
18000 & 146060 & 5 & 1224106 & 25 & 157135 & 4\\
19000 & 152703 & 5 & 1229229 & 25 & 157135 & 4\\
20000 & 163555 & 6 & 1234836 & 25 & 157135 & 3\\
\noalign{\smallskip}\hline
\label{tab:suppvartandem} 
\end{tabular}
\end{table}
